Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-30 tex/paper.tex
save

Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{Z|f(X;\theta_f)} =
p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$.

@@

For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{Z|f(X;\theta_f)} =
p_{r(f(X;\theta_f);\hat\theta_r)|f(X)}$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f$ when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the log-loss $L_f$
    reduces to $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
   -\sum_{z_l \in {\cal Z}} p(z|\smash{\hat\theta}_f) \log  p(z|\smash{\hat\theta}_f) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
\end{itemize}
Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
point, because of Proposition~\ref{prop:1}, $(\smash{\hat\theta}_f, \smash{\hat\theta}_r)$, $f$ is both an optimal
classifier and a pivotal quantity.
\end{proof}

>>>>

Let us first recall that the cross-entropy for distributions $p$ and $q$ is
minimized when $p=q$. For ${\cal L}_r$ defined as the cross-entropy between the
true conditional distribution of the nuisance $p_{Z|f(X;\theta_f)}$ and
the approximate conditional distribution of the nuisance
$p_{r(f(X;\theta_f);\theta_r)|f(X)}$, the optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$ are therefore such that $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$.

@@

For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{r(f(X;\theta_f);\hat\theta_r)|f(X)} = p_{Z|f(X;\theta_f)}$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
\log |{\cal Z}| \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the log-loss $L_f$
    reduces to $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z_l|f(X;\smash{\hat\theta}_f)) = \frac{1}{|{\cal Z}|}$ for all $z_l \in {\cal
   Z}$ and $H(p_{Z|f(X;\smash{\hat\theta}_f)}) =
   -\sum_{z_l \in {\cal Z}} p(z|\smash{\hat\theta}_f) \log  p(z|\smash{\hat\theta}_f) =  -\sum_{z_l \in {\cal Z}} \frac{1}{|{\cal Z}|} \log \frac{1}{|{\cal Z}|} = \log |{\cal Z}|$.
\end{itemize}
Accordingly, the second condition implies that $r(f(X;\smash{\hat\theta}_f) ;
\smash{\hat\theta}_r )_l = \frac{1}{|{\cal Z}|}$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}
